Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-31 TACAS2018/paper.tex
wording

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs\footnote{Notice that since we
can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully
observable.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1} and \eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
$k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$ is the
initial state.  In this work we raise the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, starting from an initial set
$X_0\subseteq \mathbb{R}^p$, with inputs taken from the set $U \subseteq
\mathbb{R}^q$, are sequences
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs
($X_0$ and $U$, respectively), and write $\tau(X_0,U)$ to denote the set of
states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$ reached from $X_0$ by $\tau$ in one step.
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states
reached from $X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear
transformations preserve convexity~\cite{boyd2004convex}, the sets $X_n =
\tau^n(X_0,U)$ are also convex.  We define the \emph{$n$-reach tube} as
%
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
%
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer
to as the reach space of the system, generalizes the previous notions over
an unbounded time horizon (transitive closure).

@@

In many models of real-world control systems, the internal state $x$ cannot
be directly measured, but is rather inferred from its effect on the measured
output $\vec{y}$ using an observer~\cite{Astrom08}.  Essentially, a state
observer is a new dynamical model (with variable denoted as $\hat{\vec{x}}$)
that provides an estimate of the internal state $x$ of the given model, from
measurements of its inputs and outputs.

A closed-loop digital control interconnection with an observer is
illustrated in Figure~\ref{fig:observersystem}.  The linear feedback
controller (bottom left) is characterized by a gain matrix $\mat{K}_{\langle
C \rangle}$ that influences the closed-loop dynamics, and is driven by the
observer state $\hat{\vec{x}}$.  Moreover, $\vec{y}_e=\hat{\vec{y}} -
\vec{y}$ is the error on the output estimator,
%
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
%
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.  Note
that the controller and the observer are implemented in a discrete domain
(quantities within this domain are identified by the index $\langle C
\rangle$), whereas the plant evolves in a continuous domain.  We~further
model the effects of using a discrete controller as the noise signals
$\nu_1$(ADC quantization), $\nu_2$ (DAC quantization), $\nu_1'$ (rounding of
the observer state error), $\nu_2'$ (rounding of the observer state), and
$\nu_t$ (output delay of the digital system) Lastly, $k_r$ is a reference
gain required by the model.  Our goal in this work is to synthesize both
$\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, so that
specific closed-loop properties of the overall interconnection are
satisfied.

>>>>

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs\footnote{Notice that since we
can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully
observable.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1} and \eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$
is the initial state.  In this work we apply the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, starting from an initial set
$X_0\subseteq \mathbb{R}^p$, with inputs taken from the set $U \subseteq
\mathbb{R}^q$, are sequences
%
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $
%
where
%
$ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k \;.
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs
($X_0$ and $U$, respectively), and write $\tau(X_0,U)$ to denote the set of
states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$ reached from $X_0$ by $\tau$ in one step.
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states
reached from $X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear
transformations preserve convexity~\cite{boyd2004convex}, the sets $X_n =
\tau^n(X_0,U)$ are also convex.  We define the \emph{$n$-reach tube} as
%
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
%
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer
to as the reach space of the system, generalizes the previous notions over
an unbounded time horizon (transitive closure).

@@

In many real-world control systems the internal state $x$ cannot be directly
measured, but is rather inferred from its effect on the measured
output~$\vec{y}$ using an observer~\cite{Astrom08}.  Essentially, a state
observer is a new dynamical model (with variable denoted as $\hat{\vec{x}}$)
that provides an estimate of the internal state $x$ of the given model using
measurements of its inputs and outputs.

A closed-loop digital control interconnection with an observer is
illustrated in Figure~\ref{fig:observersystem}.  The linear feedback
controller (bottom left) is characterized by a gain matrix $\mat{K}_{\langle
C \rangle}$ that influences the closed-loop dynamics, and is driven by the
observer state $\hat{\vec{x}}$.  Moreover, $\vec{y}_e=\hat{\vec{y}} -
\vec{y}$ is the error on the output estimator,
%
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
%
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.  Note
that the controller and the observer are implemented in a discrete domain
(quantities within this domain are identified by the index $\langle C
\rangle$), whereas the plant evolves in a continuous domain.  We~further
model the effects of using a discrete controller as the noise signals
$\nu_1$(ADC quantization), $\nu_2$ (DAC quantization), $\nu_1'$ (rounding of
the observer state error), $\nu_2'$ (rounding of the observer state), and
$\nu_t$ (output delay of the digital system).  Finally, $k_r$ is a reference
gain required by the model.  Our goal in this work is to synthesize both
$\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$ so that
specific closed-loop properties of the overall system are
satisfied.
